\chapter{Introduction}\label{chapter:Intro}

Financial regulation creates structural constraints on network formation. When regulations prohibit direct bilateral relationships, agents must route transactions through connected parties, generating inefficiency costs. Traditional policy responses focus on promoting competition among intermediaries within the constrained network. This thesis examines an alternative mechanism: parallel infrastructure operating under different regulatory treatment. This alternatives infrastructures, that could be for example, non-banking financial intermediaries (NBFI), today represent \$238 trillion globally—49.1\% of global financial assets \cite{board_global_2024}.

We demonstrate that competition among intermediaries provides only partial welfare recovery. In contrast, dual-network infrastructure achieves complete welfare restoration through coexistence rather than displacement of traditional systems.

\section{Regulatory Constraints and Network Efficiency}

Regulatory restrictions serve policy objectives—prudential regulation limits exposure to risky counterparties, capital controls manage cross-border flows, ring-fencing separates banking activities. These restrictions impose costs by preventing direct contracting when agents would benefit from bilateral relationships.

We quantify this cost in a three-agent network where regulation prohibits one direct connection. In our example, the constraint erases 50.8\% of potential welfare relative to an unconstrained benchmark. A National Bank and Local Firm can only transact through a Local Bank, creating a structural bottleneck.

The conventional response is promoting competition. Introducing a second central agent (Fintech) that competes with the Local Bank increases welfare by 46.5\% relative to the baseline. However, a 27.9\% welfare gap persists. Competition improves outcomes but cannot overcome the structural inefficiency created by the constraint itself.

\section{Alternative Network Infrastructure}

Alternative financial infrastructures operate alongside traditional systems under different regulatory treatment:

\begin{itemize}
    \item Shadow banks, or NBFI operate outside traditional banking regulations 
    \item Offshore financial centers enable cross-border transactions that domestic regulations constrain
    \item Non-bank financial institutions operate under distinct regulatory frameworks
    \item Fintech platforms offer payments and credit outside traditional banking
    \item Decentralized finance protocols enable trading and lending without centralized intermediaries
\end{itemize}

These infrastructures share a mechanism: they enable financial relationships that regulatory constraints prevent in traditional systems. Shadow banks lend to borrowers that prudential regulation restricts traditional banks from serving. Offshore centers process transactions that capital controls block domestically. DeFi protocols facilitate derivatives trading that exchange regulations prohibit.

The question is whether alternative infrastructure can restore efficiency losses from regulatory constraints while restrictions remain in the traditional network.

\section{The Dual-Network Framework}

We model agents accessing two parallel networks: a Traditional Network subject to regulatory constraints, and an Alternative Network operating under different regulatory treatment. Agents allocate contracts strategically across networks to maximize welfare, subject to a portfolio constraint $w^*_i = w^T_i + w^A_i$.

The Alternative Network absorbs contracts prohibited in the Traditional Network. The Traditional Network retains volume where established infrastructure can offer advantages. Coexistence generates aggregate welfare matching the unconstrained benchmark—complete efficiency recovery.

This mechanism differs from pure competition. Under competition alone, the Local Bank's utility falls 45.7\%. Under dual-network equilibrium, all agents experience utility gains: central agents increase 6.9\% each, while the blocked party (Local Firm) captures 84.3\% of total welfare improvement.

\section{Theoretical Framework}

Our analysis extends \citeA{jalan_incentive-aware_2024}, who characterize pairwise stable equilibria where agents negotiate bilateral contracts. We introduce three extensions:

\subsection{Dual Network Competition}

While \citeA{jalan_incentive-aware_2024} analyze equilibria in a single network, we model competition between two parallel infrastructures. The portfolio constraint $w^*_i = w^T_i + w^A_i$ fixes total exposure but allows optimized allocation between Traditional ($T$) and Alternative ($A$) networks.

This captures:
\begin{itemize}
    \item Asymmetric regulatory treatment (constraints differ across networks)
    \item Network-specific costs (frictions $\lambda_A$)
    \item Network-specific benefits (network advantages $\tau$)
\end{itemize}

\subsection{Transparency-Friction Trade-off}

Alternative infrastructures exhibit trade-offs between informational benefits and operational costs. Parameter $\tau$ captures reductions in information asymmetries—blockchain's public ledger visibility, shadow banking's credit assessment, NBFI's relationship monitoring. Parameter $\lambda_A$ captures operational costs including transaction fees, settlement delays, and liquidity fragmentation.

Empirical evidence shows cost variation: traditional cross-border remittances average 6.35\% fees, while alternative channels achieve under 1\% \cite{world_bank_remittance_2024}. Our model characterizes when these trade-offs favor adoption versus concentration in traditional systems.

\section{Main Results}

\subsection{Competition Provides Partial Recovery}

Competition among central agents reduces rents but cannot eliminate deadweight loss from regulatory constraints. In the example used, the constraint destroys 50.8\% of welfare. Adding a competing agent improves welfare, but a gap persists.

This establishes that competition is necessary but insufficient. The structural constraint creates inefficiency that competition cannot overcome.

\subsection{Alternative Infrastructure Achieves Complete Recovery}

A dual-network system where the Alternative Network operates without the constraint achieves utility of 1.43, matching the unconstrained benchmark. This represents 103\% improvement over the baseline and eliminates the 27.9\% welfare gap.

The mechanism operates through strategic coexistence: agents route 72\% of bilateral volume through the Alternative Network, retaining 28\% in Traditional channels. The prohibited National Bank-Local Firm connection migrates entirely to the Alternative Network.

\subsection{Regulatory Arbitrage Drives Adoption}

Alternative Network adoption exhibits sensitivity to market structure. Regulatory arbitrage—accessing blocked contracts—drives adoption rather than technological superiority. When competition improves Traditional Network activity, the relative advantage of alternative infrastructure declines proportionally.

\subsection{Welfare Gains Concentrate on Blocked Parties}

The Local Firm (prevented from directly contracting with the National Bank) captures 84.3\% of aggregate welfare gains. This reflects that alternative infrastructure directly addresses the constraint binding this party. Intermediaries receive 6.9\% gains each.

Unlike pure competition (which reduces incumbent utility 45.7\%), the structural solution generates positive gains for all parties.

\section{Policy Implications}

\subsection{Infrastructure Policy Over Competition Policy}

When welfare losses stem from structural constraints, enabling alternative infrastructure (0\% welfare gap) dominates promoting competition within constrained networks (27.9\% gap). Competition improves Traditional Network efficiency but cannot eliminate inefficiency from forced intermediation.

Where constraints no longer serve original objectives, removing them captures all welfare gains. Where constraints serve legitimate objectives (systemic risk mitigation, consumer protection), enabling alternative infrastructure achieves welfare restoration while maintaining Traditional Network restrictions.

\subsection{Design for Coexistence, Not Displacement}

The optimal outcome arises from coexistence. Traditional Network retains 52.1\% of volume despite the constraint. Agents use Alternative Networks for blocked transactions while maintaining Traditional relationships where established infrastructure offers operational advantages.

Policy should facilitate access to alternative infrastructure—clarifying legal treatment of shadow banking, providing regulatory sandboxes for fintech, establishing frameworks for decentralized protocols—rather than mandating migration or prohibiting parallel systems.

\subsection{Competition and Infrastructure Are Complementary}

Competition reduces Alternative Network adoption from 72.4\% to 47.9\%, indicating partial substitutability. However, both remain valuable: welfare recovery is complete despite lower adoption because competition improves Traditional Network activity under constraint.

Regulators should pursue both mechanisms: enabling alternative infrastructure access and reducing barriers to entry. The combination generates welfare superior to either approach alone.

\section{Relation to Alternative Infrastructures}

\paragraph{Shadow Banking and NBFIs}
Shadow banking emerged to provide credit outside prudential regulation. Our model predicts adoption is highest where traditional banking faces regulatory constraints—consistent with shadow banking's growth following post-2008 capital requirements \cite{buchak_fintech_2018}.

\paragraph{Offshore Financial Centers}
Cross-border flows route through offshore jurisdictions to circumvent capital controls or tax regulations. Our framework formalizes this: the offshore system serves as Alternative Network enabling transactions that domestic regulation blocks.

\paragraph{Fintech and P2P Lending}
Peer-to-peer platforms connect borrowers directly with lenders, bypassing traditional bank intermediation. Our results suggest these platforms succeed in markets where banking regulation constrains traditional lending rather than achieving technological superiority. \citeA{buchak_fintech_2018} provide empirical support: 60\% of shadow banking expansion stems from regulatory burdens.

\paragraph{Decentralized Finance Protocols}
Blockchain-based DeFi enables trading, lending, and derivatives without centralized intermediaries. Our framework predicts adoption is highest where traditional systems face regulatory constraints, transparency benefits are valuable, operational frictions remain manageable, and traditional intermediaries operate with limited competition.

This pattern aligns with observed DeFi adoption in cross-border payments (traditional fees 6-12\%) and decentralized derivatives (regulatory constraints limit traditional access).

The key insight: alternative infrastructure value stems from regulatory arbitrage—accessing prohibited contracts—rather than technological advantages alone. This reframes policy debates: rather than "should we allow alternative networks?", regulators should examine whether restrictions justify the circumvention they induce.

\section{Document Structure}

Chapter \ref{capreview} reviews literature on financial networks, platform competition, and alternative infrastructures.

Chapter \ref{capMethods} develops the theoretical model, extending \citeA{jalan_incentive-aware_2024} to incorporate dual-network competition, transparency-friction trade-offs, and asymmetric regulatory constraints.

Chapter \ref{cap:Results} presents findings through four scenarios: (1) baseline inefficiency from regulatory constraints, (2) partial recovery through competition, (3) conceptual validation of the structural solution, and (4) complete recovery through dual-network stability.

Chapter \ref{chap:conclusion} concludes with policy implications, limitations, and directions for future research.